Even-point Multi-loop Unitarity and its Applications: Exponentiation, Anomalies and Evanescence

### 2023-07-14 eventPointTwoLoop.tex
fixed some typos

\definecolor{NUpurple}{RGB}{078,042,132}

\def\dj#1{{\color{NUpurple}\it \bf JJ: #1}} 
\def\needRefs{{\color{red} [[ REFs NEEDED ]]}}

@@

\abstract{
We identify novel  structure in newly computed multi-loop amplitudes and quantum actions for even-point effective field theories, including both the nonlinear sigma model (NLSM) and double-copy gauge theories such as Born-Infeld and its supersymmetric generalizations.  
We exploit special properties of all even-point theories towards efficient unitarity based amplitude construction. 
We find evidence that the leading IR divergence of NLSM amplitudes exponentiates. 
We then systematically compute the two-loop anomalous behavior of Born-Infeld, and find that the counterterms required to restore $U(1)$ behavior at loop-level in the Born-Infeld theory can be constructed via a symmetric-structure double-copy.  
We also demonstrate that the one-minus $(-+++)$ two-loop amplitude vanishes upon the introduction of an evanescent operator. In addition to these pure photon counterterms, we verify through explicit calculation that the anomalous matrix elements that violate $U(1)$ duality invariance can be alternatively cancelled by summing over internal $\mathcal{N}=4$ DBIVA superfields. 
Finally we find that $\mathcal{N}=4$ Dirac-Born-Infeld-Volkov-Akulov (DBIVA) amplitudes permit double-copy construction through two-loop order by reproducing our unitarity based result with a double copy between color-dual $\mathcal{N}=4$ super-Yang-Mills  and our two-loop NLSM amplitudes. 
This result provides strong evidence for the existence of a color-dual representations for NLSM beyond one-loop. 
We conclude with an overview of how $D$-dimensional four-photon counterterms can be constructed in generality with the symmetric-structure double-copy, and outline a convenient way of counting evanescent operators using Hilbert series as generating functions. }

@@

The motivation for this work is three-fold. First, perturbative calculations in NLSM and DBIVA effective field theories are in many ways significantly simpler than the typical multi-loop calculations of more phenomenological theories, like quantum-chromodynamics (QCD). As we will show, the multi-loop amplitudes for these theories allow us to recycle many of the $D$-dimensional integration tools that are so powerful at one-loop order. To approach this problem, we expand on the now standard procedure of Forde \cite{Forde:2007mi}, that computes one-loop amplitudes directly from 4D unitarity cuts. In \sect{sec:EMU} we'll show that there are only two basis integrals needed at two-loop four-point for even-point effective field theory amplitudes. In the spirit of one-loop unitarity, the integral coefficients can be directly extracted from Even-point Multi-loop Unitarity and tensor reduction. Our method of Even-point Multi-loop Unitarity (EMU) is a $D$-dimensional approach to integrand construction. This allows us to compute a large catalog of $D$-dimensional two-loop amplitudes, which provides fertile ground for cultivating insights about the multi-loop structure of effective field theories. The perturbative depth of our calculations sheds light on the exponential structure of IR divergences for NLSM amplitudes through two-loops, the anomalous behavior $U(1)$ duality invariance through two-loop order in DBIVA theories, as well as the emergence of evanescent operators relevant for anomaly cancellation in pure Born-Infeld theory. We believe these surprisingly rich physical structures could serve as a theoretical laboratory for future studies of the interplay between effective field theory operators and perturbative calculations in quantum field theory.

@@

The remarkable simplicity of one-loop amplitudes is even more dramatic in a fixed spacetime dimension. In $D=4$, the kinematic projection to a basis of scalar integrals saturates at the box integral due to the appearance of Gram determinants. Roughly speaking, $N$-gon integral coefficients of \eqn{TenRed1loop} come dressed with a factor of $\mathcal{G}_N$,
\begin{equation}
\mathcal{G}_N = \text{det} (k_i \cdot k_j)
\end{equation}
where $k_i$ are the external momenta flowing into the $N$-gon integral. In fixed spacetime dimension, $D$, the momenta are $D$-component vectors and thus the Lorentz product matrix $(k_i \cdot k_j)$ must have a null space for $N>D$. As a result, $C_M \neq 0$, only when $M\leq D+1$. Furthermore, in $D=4-2\epsilon$ dimensions, any scalar pentagon integral can be rewritten up to $\mathcal{O}(\epsilon)$ in terms of five pinched scalar box integrals \cite{Bern:1993kr}. All the information of one-loop 4D amplitudes can therefore be completely determined by evaluating up to box integrals, regardless of the multiplicity. We will demonstrate in the next section how this can be used in 4D integral construction.

The last important feature of one-loop basis integrals is that many of them can be evaluated in arbitrary dimension for a very general set of parameters appearing in the exponents of the propagators. Take for example the $D$-dimensional triangle and bubble integrals \cite{Smirnov:2004ym} that we will use throughout the text:
\begin{equation}\label{eq:integrals}
{
\begin{aligned}
I_{2,(K)}^{(\alpha_1,\alpha_2)}&=\int \frac{d^{2D} l}{(2\pi)^{2D}} \frac{1}{[l^2]^{\alpha_1}[(l+K)^2]^{\alpha_2}} 
\\
&=i\left[-\frac{K^2}{4\pi}\right]^D\frac{\Gamma(D-\alpha_1)\Gamma(D-\alpha_2)\Gamma(\alpha_{12}-D)}{[K^2]^{\alpha_{12}}\Gamma(\alpha_1)\Gamma(\alpha_2)\Gamma(2D-\alpha_{12})} 
 \\\\
I_{3,(K_{12})}^{(\alpha_1,\alpha_2,\alpha_3)}&= \int \frac{d^{2D} l}{(2\pi)^{2D}} \frac{1}{[l^2]^{\alpha_1}[(l+K_1)^2]^{\alpha_2}[(l+K_{12})^2]^{\alpha_3}} 
\\
&=i\left[-\frac{K_{12}^2}{4\pi}\right]^D\frac{\Gamma(D-\alpha_{12})\Gamma(D-\alpha_{23})\Gamma(\alpha_{123}-D)}{[K_{12}^2]^{\alpha_{123}}\Gamma(\alpha_1)\Gamma(\alpha_3)\Gamma(2D-\alpha_{123})} 
\end{aligned}
}
\end{equation}
where we have introduced the notation $\alpha_{12...n} = \alpha_1+\alpha_2+\cdots +\alpha_n$, and have taken $D\rightarrow 2D$ to make the expressions more compact. The above integral expressions hold for massless external kinematics, $K_i^2=0$, which applies to all the integrals needed for the physical processes described in the text.

The existence of closed form expressions, like those in \eqn{eq:integrals}, while common for one-loop integrals, are incredibly rare for higher loop order outside of very specialized kinematic regimes. Luckily, in this work, the integral basis at four-point two-loop order for even-point theories can be reconstructed from the one-loop expressions in \eqn{eq:integrals}. This is a special property of we refer to as recursively one-loop amplitudes, which we can exemplify in the simple context of scalar $\phi^{2k}$ theory.

@@

We could also consider higher derivative generalizations of the NLSM that are invariant under the shift symmetry of \eqn{eq:shiftSym}.  Such  EFTs correspond to the so-called chiral limit of {chiral perturbation theory}~\cite{Weinberg:1978kz,Gasser:1983yg,Kampf:2006yf}, describing massless pion scattering below the chiral symmetry breaking scale of QCD:
\begin{equation}\label{eq:XPTLag}
\mathcal{L}^{\chi\text{PT}}=\frac{1}{2}\text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)] + \frac{\beta_1}{f_\pi^2} \text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)]^2 + \frac{\beta_2}{f_\pi^2} \text{tr}[\partial_\mu U^\dagger \partial_\nu U]\text{tr}[\partial_\mu U^\dagger \partial_\nu U] +\cdots
\end{equation}
By construction, each of the higher derivative operators appearing above are invariant under the shift symmetry of \eqn{eq:shiftSym}. These higher derivative generalizations of NLSM have been studied both in the context of the $S$-matrix bootstrap \cite{Manohar:2008tc,Bellazzini:2016xrt,Guerrieri:2020bto}, and soft theorem bootstraps of refs. \cite{Cheung:2014dqa,Cheung:2015ota,Cheung:2016drk,Cheung:2018oki,Low:2019ynd}. Considering the relevance of \eqn{eq:XPTLag} in Standard Model pion scattering, the amplitudes for $SU(2)$ NLSM have been computed through two-loop order \cite{Bijnens:1995yn,Bijnens:1997vq,Girlanda:1997ed}. In this work, we add to the literature by computing the two-loop amplitudes of \eqn{eq:NLSMLag} for arbitrary $SU(N_c)$ color structure. In doing so, we show that the IR divergences of the theory exponentiates in the large $N_c$ limit. 
\paragraph{Born-Infeld Photons and Dirac Scalars}
The second even-point EFT we will study is Born-Infeld theory \cite{Born:1934gh}. This theory describes the dynamics of open string endpoints with $U(1)$ charge attached to a $D$-dimensional spacetime with Dirichlet boundary conditions. Given the stringy dynamics orthogonal to the spacetime, the electric field generated by the charged endpoints has a maximum field strength of order the inverse string tension, $1/\alpha'$. The effective Lagrangian for this theory
\begin{equation}
 \label{biLag}
  \alpha'^{\,2}\mathcal{L}^{\text{BI}} = 1-\sqrt{\text{det}(\eta_{\mu\nu}+\alpha' F_{\mu\nu})} \,.
\end{equation}
A close cousin of Born-Infeld theory is the dimensionally reduction to Dirac scalars, which describes a $D$-dimensional spacetime propagating in a $(D+1)$-dimensional background:
\begin{equation}
 \label{DBI}
  \alpha'^{\,2}\mathcal{L}^{\text{DBI}} =1- \sqrt{\text{det}(\eta_{\mu\nu}+\alpha'^{\,2} \partial_\mu \varphi \partial_\nu \varphi)}
\end{equation}
where $\varphi$ is the spacetime coordinate in the orthogonal space. This theory is uniquely determined by considering the most general polynomial  of $X=(\partial \varphi)^2$, $P(X)$,  constrained to be  invariant under the field redefinition~ \cite{Cheung:2016drk}:
\begin{equation}\label{DBIShift}
\varphi \rightarrow \varphi + c + b^{\mu} (x_\mu + \varphi \partial_\mu \varphi)
\end{equation}

@@

Before proceeding, we comment briefly on the higher derivative corrections one might expect to appear in duality invariant effective field theories. In the case of supersymmetric DBIVA at four-points, $U(1)$ symmetry is promoted to an $R$-symmetry, and is thus protected perturbatively to all orders by supersymmetric Ward Identities \cite{Heydeman:2017yww}. We show this explicitly in the case of $\mathcal{N}=4$ DBIVA through two-loop in \sect{sec:DBIU}. However, this supersymmetric enhancement of $U(1)$ duality does not in general apply to higher multiplicity amplitudes. Beyond four-point $R$-symmetry permits non-vanishing matrix elements outside the split helicity sector, beyond leading order in the higher-derivative expansion \cite{Pavao:2022kog}. The low energy effective action of the OSS is an exemplar of an EFT that respects U(1) duality at leading order in $\alpha'$, but permits duality violating matrix elements at six-point and beyond \cite{Carrasco:2016ldy}. At these higher orders in $\alpha'$, while preserving $R$-symmetry, the abelian OSS violates $U(1)$ duality at tree-level. It would be fascinating if these higher derivative violations of $U(1)$ duality in OSS spectrum are precisely those needed to cancel anomalous behavior at the quantum level. We see this as a natural future direction and application of the methods we develop in this work.

@@

\paragraph{\textbf{Construction}} Using the four-point tree-level NLSM amplitude , and following the reasoning of \sect{sec:EMU}, it is straightforward to write down an off-shell integrand associated with the one-loop bubble:
\begin{equation}
\scaleIntApion =\int \frac{d^D l}{(2\pi)^D} \frac{ \sum_{\text{states}} \mathcal{A}^{\text{NLSM}}_{(\bar{l}_1|12|\bar{l}_2)}\mathcal{A}^{\text{NLSM}}_{(l_2|34|l_1)}}{l^2(l+k_{12})^2}\,.
\end{equation}
We define the internal loop momenta as, $\bar{l}_1 =-l_1= l$ and $\bar{l}_2 =-l_2= -(l+k_{12})$. The integrand numerator can be easily evaluated, yielding the following expression:
\begin{align}
  \sum_{\text{states}} \mathcal{A}^{\text{NLSM}}_{(\bar{l}_1|12|\bar{l}_2)}\mathcal{A}^{\text{NLSM}}_{(l_2|34|l_1)}&= 4f_\pi^{-4}C^{\text{box}}_{(1234)} \left[(k_2\cdot \bar{l}_1)(k_3\cdot l_1)+(k_1\cdot \bar{l}_1)(k_4\cdot l_1)\right]\\
&+4f_\pi^{-4}C^{\text{box}}_{(1243)} \left[(k_1\cdot \bar{l}_1)(k_2\cdot \bar{l}_1)+(k_3\cdot l_1)(k_4\cdot l_1)\right]
\end{align}
Again, as following the reasoning in \sect{sec:EMU} we do not need to worry about any contact terms that may have been missed in this cut -- any such terms do not exist for an even-point theory.

@@

Now, given the appearance of loop momenta in the integrand, we must proceed to the next step in our integration procedure, tensor reduction. 

@@

A^{4-2\epsilon}_{(1234)} &=\frac{i}{12\pi^2 }f_{\pi}^{-4}\left[\frac{\sigma_2}{\epsilon}+\left(s_{12}(s_{13}-s_{12})\frac{\text{ln}(-s_{12})}{8}+(1\leftrightarrow 3)\right)\right]+\mathcal{O}(\epsilon)

@@

A^{2-2\epsilon}_{(1234)} &= -\frac{i}{2\pi }f_{\pi}^{-4}\left[\frac{s_{13}}{\epsilon}-s_{13}\frac{\text{ln}(-s_{12})+\text{ln}(-s_{23})-3}{2}\right] +\mathcal{O}(\epsilon) \label{eq:1loopPionD2}

@@

where $\sigma_2$ is the previously defined four-point permutation invariant, and the full amplitudes are recovered by summing over cyclic permutations of (2,3,4), 

@@

It's worth pointing out that while the divergence in $D=4-2\epsilon$ is an ultraviolet divergence, the one in $D=2-2\epsilon$ is a logarithmic IR divergence, akin to that of $\mathcal{N}=4$ super-Yang-Mills (sYM) in the critical dimension of $D=4-2\epsilon$. Since all the particle states above are scalars, there are no helicity structures to map into, and thus we will bypass the final integration step of \textbf{projection}. Equipped with this one-loop warmup, we are now prepared to take on a two-loop example. 

@@

\scaleIntBscalarsmall{1}{2}{3}{4} &= f_{\pi}^{-6}C^{\text{2box}}_{(12|34)} \frac{s_{12}}{6} \left[\frac{(D-1)(D-4)s_{14}+2(D-2)^2 s_{13}}{(D-1)(4-3D)}\right][I_3\circ I_2]^D(k_{12})

@@

&+f_{\pi}^{-6}C^{\text{2box}}_{(12|43)} \frac{s_{12}}{6} \left[\frac{(D-1)(D-4)s_{13}+2(D-2)^2 s_{14}}{(D-1)(4-3D)}\right][I_3\circ I_2]^D(k_{12})

@@

&+f_{\pi}^{-6}C^{\text{Xbox}}_{([12]|34)}\left[\frac{s_{12}^2}{6}\frac{D+1}{D-1}\right][I_3\circ I_2]^D(k_{12})

@@

[I_3\circ I_2]^D(k_{12}) \equiv \int \frac{d^Dl_1}{(2\pi)^D} \frac{d^Dl_2}{(2\pi)^D} \frac{(l_2+k_1)^2}{l_1^2  l_2^2 (l_1+l_2+k_1)^2(l_2+k_{12})^2}\,.

@@

C^{\text{2box}}_{(12|34)} \xrightarrow{N_c\rightarrow\infty } N_c^2\, C(1234) +\mathcal{O}(N_c)

@@

where we have defined the color trace factor, $C(1234)\equiv  \text{tr}[T^1T^2 T^3 T^4]$. Furthermore, the leading divergences for the loop integrals in $D=2-2\epsilon$ dimensions are the following:

@@

The analogous divergences for $D=4-2\epsilon$ can be extracted directly from the integral expressions in \eqn{eq:integrals}. Plugging these values into \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich}, we can extract the full leading logarithmic divergence from the full color dressed amplitude in \eqn{eq:NLSM2loop}. Since we are interested in the large $N_c$ limit, we will define a color-ordered quantity $A^{2-2\epsilon}_{(1234)}$, 

@@

A^{2-2\epsilon}_{(1234)} = -\frac{1}{2}\left[\frac{f_{\pi}^{-2}N_c }{4\pi \epsilon}\right]^2 \,f_\pi^{-2}s_{13} 

@@

It would be interesting to investigate whether this exponential structure also can be applied to the subleading logarithms in the planar limit, similar to what was found in \cite{Anastasiou:2003kj,Bern:2005iz} in $D=4-2\epsilon$ for planar $\mathcal{N}=4$ sYM. The iterated structure of planar $\mathcal{N}=4$ has been linked to the integrability of the theory \cite{Sterman:2002qn,Minahan:2002ve,Bena:2003wd,Beisert:2003jj,Beisert:2003jb,Beisert:2003yb,Dolan:2003uh,Arutyunov:2003rg,Ryzhov:2004nz,Frolov:2005iq}. We find compelling the possibility that this iterated structure of NLSM amplittudes in $D=2-2\epsilon$ could indicate similar integrable behavior, as studied in \cite{Shankar:1977cm,Zamolodchikov:1977nu,Zamolodchikov:1978xm, Komatsu:2019hgc}. We leave this as a direction of future study.

@@

&\frac{(D^3- 24 D^2+ 68 D +72 )}{8}f_{12}f_{34}

@@

&+\frac{(2D^2-3D-8 )}{2}(f_{1234} + f_{1243}) 

@@

>>>>

\definecolor{NUpurple}{RGB}{078,042,132}
\def\dj#1{{\color{NUpurple}\it \bf JJ: #1}} 
\def\dn#1{{\color{funRed}\it \bf NHP: #1}} 
\def\needRefs{{\color{red} [[ REFs NEEDED ]]}}

@@

\abstract{
We identify novel  structure in newly computed multi-loop amplitudes and quantum actions for even-point effective field theories, including both the nonlinear sigma model (NLSM) and double-copy gauge theories such as Born-Infeld and its supersymmetric generalizations.  
We exploit special properties of all even-point theories towards efficient unitarity based amplitude construction. 
We find evidence that the leading IR divergence of NLSM amplitudes exponentiates when the symmetry group is $SU(2)/U(1)\cong \mathbb{CP}^1$. 
We then systematically compute the two-loop anomalous behavior of Born-Infeld, and find that the counterterms required to restore $U(1)$ behavior at loop-level in the Born-Infeld theory can be constructed via a symmetric-structure double-copy.  
We also demonstrate that the one-minus $(-+++)$ two-loop amplitude vanishes upon the introduction of an evanescent operator. In addition to these pure photon counterterms, we verify through explicit calculation that the anomalous matrix elements that violate $U(1)$ duality invariance can be alternatively cancelled by summing over internal $\mathcal{N}=4$ DBIVA superfields. 
Finally we find that $\mathcal{N}=4$ Dirac-Born-Infeld-Volkov-Akulov (DBIVA) amplitudes permit double-copy construction through two-loop order by reproducing our unitarity based result with a double copy between color-dual $\mathcal{N}=4$ super-Yang-Mills  and our two-loop NLSM amplitudes. 
This result provides strong evidence for the existence of a color-dual representations for NLSM beyond one-loop. 
We conclude with an overview of how $D$-dimensional four-photon counterterms can be constructed in generality with the symmetric-structure double-copy, and outline a convenient way of counting evanescent operators using Hilbert series as generating functions. }

@@

The motivation for this work is three-fold. First, perturbative calculations in NLSM and DBIVA effective field theories are in many ways significantly simpler than the typical multi-loop calculations of more phenomenological theories, like quantum-chromodynamics (QCD). As we will show, the multi-loop amplitudes for these theories allow us to recycle many of the $D$-dimensional integration tools that are so powerful at one-loop order. To approach this problem, we expand on the now standard procedure of Forde \cite{Forde:2007mi}, that computes one-loop amplitudes directly from 4D unitarity cuts. In \sect{sec:EMU} we'll show that there are only two basis integrals needed at two-loop four-point for even-point effective field theory amplitudes. In the spirit of one-loop unitarity, the integral coefficients can be directly extracted from Even-point Multi-loop Unitarity (EMU) and tensor reduction. Our method of EMU is a $D$-dimensional approach to integrand construction. This allows us to compute a large catalog of $D$-dimensional two-loop amplitudes, which provides fertile ground for cultivating insights about the multi-loop structure of effective field theories. The perturbative depth of our calculations sheds light on the exponential structure of IR divergences for NLSM amplitudes through two-loops, the anomalous behavior $U(1)$ duality invariance through two-loop order in DBIVA theories, as well as the emergence of evanescent operators relevant for anomaly cancellation in pure Born-Infeld theory. We believe these surprisingly rich physical structures could serve as a theoretical laboratory for future studies of the interplay between effective field theory operators and perturbative calculations in quantum field theory.

@@

The remarkable simplicity of one-loop amplitudes is even more dramatic in a fixed spacetime dimension. In $D=4$, the kinematic projection to a basis of scalar integrals saturates at the box integral due to the appearance of Gramm determinants. Roughly speaking, $N$-gon integral coefficients of \eqn{TenRed1loop} come dressed with a factor of $\mathcal{G}_N$,
\begin{equation}
\mathcal{G}_N = \text{det} (k_i \cdot k_j)
\end{equation}
where $k_i$ are the external momenta flowing into the $N$-gon integral. In fixed spacetime dimension, $D$, the momenta are $D$-component vectors and thus the Lorentz product matrix $(k_i \cdot k_j)$ must have a null space for $N>D$. As a result, $C_M \neq 0$, only when $M\leq D+1$. Furthermore, in $D=4-2\epsilon$ dimensions, any scalar pentagon integral can be rewritten up to $\mathcal{O}(\epsilon)$ in terms of five pinched scalar box integrals \cite{Bern:1993kr}. All the information of one-loop 4D amplitudes can therefore be completely determined by evaluating up to box integrals, regardless of the multiplicity. We will demonstrate in the next section how this can be used in 4D integral construction.

The last important feature of one-loop basis integrals is that many of them can be evaluated in arbitrary dimension for a very general set of parameters appearing in the exponents of the propagators. Take for example the $D$-dimensional triangle and bubble integrals \cite{Smirnov:2004ym} that we will use throughout the text:
\begin{equation}\label{eq:integrals}
{
\begin{aligned}
I_{2,(K)}^{(\alpha_1,\alpha_2)}&=\int \frac{d^{D} l}{(2\pi)^{D}} \frac{1}{[l^2]^{\alpha_1}[(l+K)^2]^{\alpha_2}} 
\\
&=i\left[-\frac{K^2}{4\pi}\right]^{D/2}\frac{\Gamma(D/2-\alpha_1)\Gamma(D/2-\alpha_2)\Gamma(\alpha_{12}-D/2)}{[K^2]^{\alpha_{12}}\Gamma(\alpha_1)\Gamma(\alpha_2)\Gamma(D-\alpha_{12})} 
 \\\\
I_{3,(K_{12})}^{(\alpha_1,\alpha_2,\alpha_3)}&= \int \frac{d^{D} l}{(2\pi)^{D}} \frac{1}{[l^2]^{\alpha_1}[(l+K_1)^2]^{\alpha_2}[(l+K_{12})^2]^{\alpha_3}} 
\\
&=i\left[-\frac{K_{12}^2}{4\pi}\right]^{D/2}\frac{\Gamma(D/2-\alpha_{12})\Gamma(D/2-\alpha_{23})\Gamma(\alpha_{123}-D/2)}{[K_{12}^2]^{\alpha_{123}}\Gamma(\alpha_1)\Gamma(\alpha_3)\Gamma(D-\alpha_{123})} 
\end{aligned}
}
\end{equation}
where we have introduced the notation $\alpha_{12...n} = \alpha_1+\alpha_2+\cdots +\alpha_n$. Throughout the text we will suppress the regularization scale that appears in the argument of the logarithms -- they an be restored by dimensional analysis. The above integral expressions hold for massless external kinematics, $K_i^2=0$, which applies to all the integrals needed for the physical processes described in the text.

The existence of closed form expressions, like those in \eqn{eq:integrals}, while common for one-loop integrals, are incredibly rare for higher loop order outside of very specialized kinematic regimes. Luckily, in this work, the integral basis at four-point two-loop order for even-point theories can be reconstructed from the one-loop expressions in \eqn{eq:integrals}. This is a special property of what we refer to as recursively one-loop amplitudes, which we can exemplify in the simple context of scalar $\varphi^{2k}$ theory.

@@

We could also consider higher derivative generalizations of the NLSM that are invariant under the shift symmetry of \eqn{eq:shiftSym}.  Such  EFTs correspond to the so-called chiral limit of {chiral perturbation theory}~\cite{Weinberg:1978kz,Gasser:1983yg,Kampf:2006yf}, describing massless pion scattering below the chiral symmetry breaking scale of QCD:
\begin{equation}\label{eq:XPTLag}
\mathcal{L}^{\chi\text{PT}}=\frac{1}{2}\text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)] + \frac{\beta_1}{f_\pi^2} \text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)]^2 + \frac{\beta_2}{f_\pi^2} \text{tr}[(\partial_\mu U)^\dagger( \partial_\nu U)]\text{tr}[(\partial_\mu U)^\dagger( \partial_\nu U)] +\cdots
\end{equation}
By construction, each of the higher derivative operators appearing above are invariant under the shift symmetry of \eqn{eq:shiftSym}. These higher derivative generalizations of NLSM have been studied both in the context of the $S$-matrix bootstrap \cite{Manohar:2008tc,Bellazzini:2016xrt,Guerrieri:2020bto}, and soft theorem bootstraps of refs. \cite{Cheung:2014dqa,Cheung:2015ota,Cheung:2016drk,Cheung:2018oki,Low:2019ynd}. Considering the relevance of \eqn{eq:XPTLag} in Standard Model pion scattering, the amplitudes for $SU(2)$ NLSM have been computed through two-loop order \cite{Bijnens:1995yn,Bijnens:1997vq,Girlanda:1997ed}. In this work, we add to the literature by computing the two-loop amplitudes of \eqn{eq:NLSMLag} for arbitrary $SU(N_c)$ color structure. In doing so, we show that the IR divergences of the theory exponentiates in the large $N_c$ limit. 
\paragraph{Born-Infeld Photons and Dirac Scalars}
The second even-point EFT we will study is Born-Infeld theory \cite{Born:1934gh}. This theory describes the dynamics of open string endpoints with $U(1)$ charge attached to a $D$-dimensional spacetime with Dirichlet boundary conditions. Given the stringy dynamics orthogonal to the spacetime, the electric field generated by the charged endpoints has a maximum field strength of order the inverse string tension, $1/\alpha'$. The effective Lagrangian for this theory
\begin{equation}
 \label{biLag}
  \alpha'^{\,2}\mathcal{L}^{\text{BI}} = 1-\sqrt{\text{det}(\eta_{\mu\nu}+\alpha' F_{\mu\nu})} \,.
\end{equation}
A close cousin of Born-Infeld theory is the dimensionally reduction to Dirac scalars, which describes a $D$-dimensional spacetime propagating in a $(D+1)$-dimensional background:
\begin{equation}
 \label{DBI}
  \alpha'^{\,2}\mathcal{L}^{\text{DBI}} =1- \sqrt{\text{det}(\eta_{\mu\nu}+\alpha'^{\,2} \partial_\mu \varphi \partial_\nu \varphi)}
\end{equation}
where $\varphi$ is the spacetime coordinate in the orthogonal space. This theory is uniquely determined by considering the most general polynomial  of $X=(\partial \varphi)^2$, $P(X)$,  constrained to be invariant under the field redefinition \cite{Cheung:2016drk}:
\begin{equation}\label{DBIShift}
\varphi \rightarrow \varphi + c + b^{\mu} (x_\mu + \varphi \partial_\mu \varphi)
\end{equation}

@@

Before proceeding, we comment briefly on the higher derivative corrections one might expect to appear in duality invariant effective field theories. In the case of supersymmetric DBIVA at four-points, $U(1)$ symmetry is promoted to an $R$-symmetry, and is thus protected perturbatively to all orders by supersymmetric Ward Identities \cite{Heydeman:2017yww}. We show this explicitly in the case of $\mathcal{N}=4$ DBIVA through two-loop in \sect{sec:DBIU}. However, this supersymmetric enhancement of $U(1)$ duality does not in general apply to higher multiplicity amplitudes. While leading order in $\alpha'$ must be duality invariant at all multiplicity for DBIVA \cite{Pavao:2022kog}, beyond leading order $R$-symmetry permits non-vanishing matrix elements outside the split helicity sector. The low energy effective action of the OSS is one such exemplar of an EFT that respects $U(1)$ duality at leading order in $\alpha'$, but produces duality violating matrix elements at higher-orders above four-point \cite{Carrasco:2016ldy}. It would be fascinating if these higher derivative violations of $U(1)$ duality in OSS spectrum are precisely those needed to cancel anomalous behavior at the quantum level. To test this would require computing a six-point one-loop in supersymmetric DBIVA. Such a calculation could in principle be performed with the double copy using the color-dual integrands recently constructed in ref.~\cite{Edison:2022jln}. We see this as a natural future direction and application of the methods we develop in this work.

@@

\paragraph{\textbf{Construction}} Using the four-point tree-level NLSM amplitude , and following the reasoning of \sect{sec:EMU}, it is straightforward to write down an off-shell integrand associated with the one-loop bubble:
\begin{equation}
\scaleIntApion =\sum_{\text{states}}\int \frac{d^D l}{(2\pi)^D} \frac{ \mathcal{A}^{\text{NLSM}}_{(\bar{l}_1|12|\bar{l}_2)}\mathcal{A}^{\text{NLSM}}_{(l_2|34|l_1)}}{l^2(l+k_{12})^2}\,.
\end{equation}
We define the internal loop momenta as, $\bar{l}_1 =-l_1= l$ and $\bar{l}_2 =-l_2= -(l+k_{12})$. The integrand numerator can be easily evaluated, yielding the following expression:
\begin{align}
  \sum_{\text{states}} \mathcal{A}^{\text{NLSM}}_{(\bar{l}_1|12|\bar{l}_2)}\mathcal{A}^{\text{NLSM}}_{(l_2|34|l_1)}&= 4f_\pi^{-4}C^{\text{box}}_{(1234)} \left[(k_2\cdot \bar{l}_1)(k_3\cdot l_1)+(k_1\cdot \bar{l}_1)(k_4\cdot l_1)\right]\\
&+4f_\pi^{-4}C^{\text{box}}_{(1243)} \left[(k_1\cdot \bar{l}_1)(k_2\cdot \bar{l}_1)+(k_3\cdot l_1)(k_4\cdot l_1)\right]
\end{align}
Again, as following the reasoning in \sect{sec:EMU} we do not need to worry about any contact terms that may have been missed in this cut -- any such terms do not exist for an even-point theory.

@@

Now, given the appearance of loop momenta in the integrand, we must proceed to the next step in our integration procedure: tensor reduction. 

@@

A^{4-2\epsilon}_{(1234)} &=\frac{i}{48\pi^2 }f_{\pi}^{-4}\left[\frac{\sigma_2}{\epsilon}+\left(s_{12}(s_{13}-s_{12})\frac{\text{ln}(-s_{12})}{2}+(1\leftrightarrow 3)\right)+\frac{1}{6}(s_{13}^2 + 2s_{12}s_{23})\right]+\mathcal{O}(\epsilon)

@@

A^{2-2\epsilon}_{(1234)} &= \frac{i}{2\pi }f_{\pi}^{-4}\left[\frac{s_{13}}{\epsilon}-s_{13}\frac{\text{ln}(-s_{12})+\text{ln}(-s_{23})-3}{2}\right] +\mathcal{O}(\epsilon) \label{eq:1loopPionD2}

@@

where $\sigma_2=s_{12}^2+s_{12}s_{23}+s_{23}^2$, and the full amplitudes are recovered by summing over cyclic permutations of (2,3,4), 

@@

It's worth pointing out that while the divergence in $D=4-2\epsilon$ is an ultraviolet divergence, the one in $D=2-2\epsilon$ is a logarithmic IR divergence, akin to that of $\mathcal{N}=4$ super-Yang-Mills (sYM) in the critical dimension of $D=4-2\epsilon$. Since all the particle states above are scalars, there are no helicity structures to map into, and thus we will bypass the final integration step of {projection}. Equipped with this one-loop warmup, we are now prepared to take on a two-loop example. 

@@

\scaleIntBscalarsmall{1}{2}{3}{4} &= f_{\pi}^{-6}C^{\text{2box}}_{(12|34)} \frac{s_{12}}{3} \left[\frac{(D-1)(D-4)s_{14}+2(D-2)^2 s_{13}}{(D-1)(4-3D)}\right][I_3\circ I_2]^D(k_{12})

@@

&+f_{\pi}^{-6}C^{\text{2box}}_{(12|43)} \frac{s_{12}}{3} \left[\frac{(D-1)(D-4)s_{13}+2(D-2)^2 s_{14}}{(D-1)(4-3D)}\right][I_3\circ I_2]^D(k_{12})

@@

&+f_{\pi}^{-6}C^{\text{Xbox}}_{([12]|34)}\left[\frac{s_{12}^2}{3}\frac{D+1}{D-1}\right][I_3\circ I_2]^D(k_{12})

@@

[I_3\circ I_2]^D(k_{12}) \equiv \int \frac{d^Dl_1}{(2\pi)^D} \frac{d^Dl_2}{(2\pi)^D} \frac{(l_2+k_1)^2}{l_1^2  (l_1+l_2+k_1)^2l_2^2 (l_2+k_{12})^2}\,.

@@

C^{\text{2box}}_{(12|34)} \xrightarrow{N_c\rightarrow\infty } N_c^2\, [C(1234)+C(1432)] +\mathcal{O}(N_c)

@@

where we have defined the color trace factor, $C(abcd)\equiv  \text{tr}[T^aT^b T^c T^d]$. Furthermore, the leading divergences for the loop integrals in $D=2-2\epsilon$ dimensions are the following:

@@

The analogous divergences for $D=4-2\epsilon$ can be extracted directly from the integral expressions in \eqn{eq:integrals}. Plugging these values into \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich}, we can extract the full leading logarithmic divergence from the color dressed amplitude in \eqn{eq:NLSM2loop}. Since we are interested in the large $N_c$ limit, we will define a color-ordered quantity $A^{2-2\epsilon}_{(1234)}$, 

@@

A^{2-2\epsilon}_{(1234)} = -\frac{1}{2}\left[\frac{f_{\pi}^{-2}N_c }{2\pi \epsilon}\right]^2 \,f_\pi^{-2}s_{13} 

@@

It would be interesting to investigate whether this exponential structure also can be applied to the subleading logarithms in the planar limit, similar to what was found in \cite{Anastasiou:2003kj,Bern:2005iz} in $D=4-2\epsilon$ for planar $\mathcal{N}=4$ sYM. The iterated structure of planar $\mathcal{N}=4$ has been linked to the integrability of the theory \cite{Sterman:2002qn,Minahan:2002ve,Bena:2003wd,Beisert:2003jj,Beisert:2003jb,Beisert:2003yb,Dolan:2003uh,Arutyunov:2003rg,Ryzhov:2004nz,Frolov:2005iq}. We find compelling the possibility that this iterated structure of NLSM amplitudes in $D=2-2\epsilon$ could indicate similar integrable behavior, as studied in \cite{Shankar:1977cm,Zamolodchikov:1977nu,Zamolodchikov:1978xm, Komatsu:2019hgc}. We leave this as a direction of future study.

@@

&\frac{(D^3- 24 D^2+ 68 D +72 )}{4}f_{12}f_{34}

@@

&+(2D^2-3D-8 )(f_{1234} + f_{1243}) 

@@

\end{align}v
